# Setting up a Research Environment {-}

Before starting on our research journey, let's make sure all of the required
software is present. This week, we will introduce different technologies that
require the following software to be present. 

**Note:** Each research project may require additional software than what is
listed here. 

In particular, please make sure to download and install: 

-   **R**: R is a free software environment for statistical computing and graphics.
    Please make sure to run the latest version [**R version 4.1.0 (Camp Pontanezen)**](https://cloud.r-project.org/)
    released on **2021-05-18**.
    
| Title         | Video Link      | Download Link   |
| ------------- | -------------   | ---------------|
| Windows: Downloading R  | [YouTube]() | [Software](https://cloud.r-project.org/bin/windows/base/) |
| macOS: Downloading R    | [YouTube]() | [Software](https://cloud.r-project.org/bin/macosx/) |


-   **RStudio**: RStudio is a free and open source integrated development environment (IDE) for R.
    There are other options to use for writing R code such as
    [Visual Studio Code](https://code.visualstudio.com/) (with [R Support](https://marketplace.visualstudio.com/items?itemName=Ikuyadeu.r)),
    R GUI, 
    [GNU Emacs](https://www.gnu.org/software/emacs/) (with [Emacs Speaks Statistics](https://ess.r-project.org/)), or even a plain text editor.
    We've opted to RStudio since it is a feature-rich program for writing R code
    and, perhaps more importantly, used heavily in both academia and industry.
    In particular, we'll be using **RStudio Desktop** 1.4.1106. 
    
| Title         | Video Link      | Download Link   |
| ------------- | -------------   | ---------------|
| Windows: Downloading RStudio  | [YouTube](https://www.youtube.com/watch?v=Zq8oxC2Rq2Q) | [Software](https://www.rstudio.com/products/rstudio/download/#download) |
| macOS:   Downloading RStudio  | [YouTube]() | [Software](https://www.rstudio.com/products/rstudio/download/#download) |

-   **R packages**: R Packages are an extension to the R language that
    provide additional features. We'll be using quite a few to tackle different
    research problems. You can install the core set of packages by opening
    RStudio and typing -- or copying and pasting -- into the console:

    ```{r, echo = FALSE, cache = FALSE}
    description_pkgs = desc::desc_get_deps()
    ordered_pkgnames = sort(description_pkgs$package[description_pkgs$type == "Imports"])
    wrapped_pkgnames = strwrap(paste(encodeString(ordered_pkgnames, quote = '"'), collapse = ", "), exdent = 1)
    install_command = paste0(
      "install.packages(c(\n ", 
      paste(wrapped_pkgnames, "\n", collapse = ""), 
      "))"
    )
    ```

    ```{r code = install_command, eval = FALSE}
    ```


-   **Compiler Toolchain** (optional): Compiler Toolchains provide the opportunity
    to construct R packages from source code and translate code into a low-level machine language called assembly.
    
| Title         | Video Link      | Download Link   |
| ------------- | -------------   | ---------------|
| Windows: Downloading Rtools  | [YouTube](https://www.youtube.com/watch?v=njZBf34Akgo) | [Software](https://cran.r-project.org/bin/windows/Rtools/) |
| macOS: Install Xcode CLI    | [YouTube](https://www.youtube.com/watch?v=Z01lzHNrSdU) | - |
| macOS: Install gfortran 8.2    | - | [`gfortran8.2`](https://github.com/fxcoudert/gfortran-for-macOS/releases/download/8.2/gfortran-8.2-Mojave.dmg) |

